Grouping and summarizing

summarizing the median life expectancy

```{r}
library(gapminder)
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))

#Filter for 1957 then summarize the median life expectancy
gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp))

gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```

Summarizing by year

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```
Summarizing by continent

```{r}
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap)) %>%
  arrange(desc(maxGdpPercap))
```
Summarizing by continent and year

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```
Line plots
Visualizing median life expectancy over time

```{r}

by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap=max(gdpPercap))

by_year

ggplot(by_year, aes(x=year, y=medianGdpPercap)) +
  geom_line()
```
```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_year_continent

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color=continent)) +
  geom_line()
```

Bar Plots

Visualizing median GDP per capita by continent

```{r}
by_continent <- gapminder %>%
  filter(year==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_continent

ggplot(by_continent, aes(x=continent, y=medianGdpPercap)) +
  geom_col()
```

Visualizing GDP per capita by country in Oceania

```{r}
oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year == 1952)

oceania_1952
ggplot(oceania_1952, aes(x=country, y=gdpPercap)) +
  geom_col()
```

Histogram

Visualizing population

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
gapminder

gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil = pop/1000000)
  
ggplot(gapminder_1952, aes(x=pop_by_mil)) +
geom_histogram(bins = 50)

```
Visualizing population with x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop)) +
  geom_histogram(bins = 50) +
    scale_x_log10()
```
Boxplots

comparing GDP per capita across continents

```{r}
gapminder_1952 <- gapminder_1952 %>%
  mutate(continente = continent)

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() +
  ggtitle("Comparing GDP per capita across continents")
```
Data Manipulation with dplyr

```{r}
library(tidyverse)
counties <- read_csv('counties.csv')

glimpse(counties)

counties
```
Selecting columns

```{r}

counties %>%
  select(state, county, population, men, women)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  arrange(desc(public_work))

counties_selected <- counties %>%
  select(state, county, population)

counties_selected %>%
  filter(population > 1000000)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  filter(state=="Texas", population>10000) %>%
  arrange(desc(private_work))
```
Select, mutate, filter and arrange

```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))
```
The verb count

```{r}
selected_counties <- counties %>%
  select(state, region, county, population, citizens)

counties %>%
  count(region, sort = TRUE) %>%
  arrange(desc(n))
```
Counting citizens by state

```{r}
county_selected <- counties %>%
  select(county, region, state, population, citizens)

county_selected %>%
  count(state, wt=citizens) %>%
  arrange(desc(n))
```
Mutating and counting

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)

counties_selected %>%
  mutate(population_walk = population*walk/100) %>%
  count(state, wt=population_walk, sort=TRUE)

```

Summarizing

```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)

counties_selected %>%
  summarize(min_population=min(population), max_unemployment=max(unemployment),  averagen_income=mean(income))
```
Summarizing by state

```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarize(total_area=sum(land_area), total_population=sum(population)) %>%
  mutate(density=total_population/total_area) %>%
  arrange(desc(density))
  
counties_selected <- counties %>%
  select(region, state, county, population)

counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop= sum(population)) %>%
  summarize(average_pop=mean(total_pop), median_pop=median(total_pop))

```

Top_n verb

```{r}
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)

counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)

counties_selected <- counties %>%
  select(region, state, county, population, income)

counties_selected %>%
  group_by(region, state) %>%
  summarize(average_income=mean(income)) %>%
  top_n(1,average_income)
```

```{r}
counties_selected <- counties %>%
  select(state, metro, population)

counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop=sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>%
  count(metro)


```
Selecting columns

```{r}
counties %>%
  select(state, county, population, professional:production) %>%
  arrange(desc(service))

counties %>%
  select(state, county, population, ends_With("work")) %>%
  filter(public_work >= 50)

counties%>%
  select(state, county, population, starts_with("income"))

counties %>%
  select(state, county, population, contains("at"))
```

Renaming a column after count

```{r}
counties %>%
  count(state)
  rename(num_counties = n)
  
counties %%
  select(state, county, povert_rate = povety)
```
Using transmute

```{r}
counties %>%
  transmute(state, county, population, density=population/land_area)
  filter(population >= 1000000) %>%
    arrante(desc(density))
```

Choosing among the four verbs

```{r}
#Change the name of the unemployment column

conutie %>%
  rename(unemployment_rate=unemployment)

#Keep the state and county columns, and the columns containing poverty

counties %>%
  select(state, county, contains("poverty"))

# Calculate the franction_women column without dropping the other columns

counties %>%
  mutate(fraction_women = women/population)

# Keep only the state, county, and employment_rate columns

counties %>%
  transmute(state, county, employment_rate = employed/population)
```

